Measuring and optimizing machine learning interpretability

### 2019-06-26 2019-reisensburg/2019-reisensburg-molnar.Rmd
start reisensburg slides

---
title: "Measuring and optimizing machine learning interpretability"
author: "Christoph Molnar, Giuseppe Casalicchio, Bernd Bischl"
date: "Sommerklausur 2019-05-24"
institute: "LMU Munich"
output: 
  beamer_presentation:
    includes:
      in_header: "header_pagenrs.tex"
---

@@

- Why measure interpretability?
- Functional Decomposition with Accumulated Local Effects
- Measures
- Application

@@

```{r, echo = FALSE}
knitr::include_graphics("../images/black-box.png")
```
## Solution: Interpretable Models?

$\Rightarrow$ Pay with loss in performance; Unclear how to compare models

```{r, echo = FALSE}
knitr::include_graphics("../images/white-box.png")
```
## Solution: Post-hoc Interpretation?

$\Rightarrow$ Post-hoc interpretation (e.g. feature importance, partial dependence plots) works better for less complex models.

@@

## Is Interpretability Unscientific?

```{r}
knitr::include_graphics("../images/spirit-board.jpg")
```

@@

$\Rightarrow$ Makes claims of interpretability more explicit ("Model A uses less features than B and has less interactions")

@@

```{r}
knitr::include_graphics("../images/perf-interpret-tradeoff-1.jpeg")
```

@@

- Model-agnostic interpretability measures: number of features, interaction strength, main effect complexity
- Use: model comparison and optimization metric

>>>>

---
title: "Measuring and optimizing machine learning interpretability"
author: "Christoph Molnar, Giuseppe Casalicchio, Bernd Bischl"
date: "Statistical Computing 2019-07-03"
institute: "LMU Munich"
output: 
  beamer_presentation:
    includes:
      in_header: "header_pagenrs.tex"
---

@@

- Minimizing model complexity maximizes interpretability
- Functional Decomposition
- Measures
- Application

@@

```{r, echo = FALSE}
knitr::include_graphics("../images/black-box.png")
```

## Solution: Post-hoc Interpretation?

<!-- $\Rightarrow$ Post-hoc interpretation (e.g. feature importance, partial dependence plots) works better for less complex models. -->

@@

## Sparsity Improves Interpretation

TODO: Image of 3 vs 100 feature in importance

## Additivity Improves Interpretation

TODO: Image from paper with PDP and increasing interactions

## Linearity Improves Interpretation

TODO: Image with increasing MEC

@@

$\Rightarrow$ Allows choosing models with improved post-hoc interpretation

@@

@@

- Model-agnostic complexity measures: number of features, interaction strength, main effect complexity
- Use: model comparison and optimization metric
